---
title: "Outline Of Analysis"
---

```{r set-knitr-opts, echo=FALSE, child = '../workflow/modules/setup_chunk.Rmd'}
```


```{r packages}
library(tidyverse)
library(DiagrammeR)
library(pander)
library(yaml)
library(scales)
```

```{r remotes, results='hide'}
## Deal with github packages
if (!"spBioUtils" %in% rownames(installed.packages()))
  BiocManager::install("steveped/spBioUtils")
stopifnot(library(spBioUtils, logical.return = TRUE))
```

```{r config}
config <- here::here("config/config.yml") %>%
  read_yaml()
samples <- config$samples %>%
  here::here() %>%
  read_tsv()
```


## Description

This is a standardised workflow for beginning the comparison between two or ChIP targets, using BAM files as the starting point.
Treatment groups and targets are specified using `config/config.yml`.

If H3K27ac is provided as a ChIP target with exactly two conditions, Super-Enhancers will be detected using the ROSE algorithm.

## Workflow

```{r plot-workflow, fig.height = 10, fig.width = 10, fig.cap = "*Summary of processing workflow. The primary data pipeline is shown in red, preparatory steps are shown in blue whilst collation of final output is in green.*"}
here::here("workflow", "rules", "rulegraph.dot") %>%
  readLines() %>%
  rm_dot_node(node = "\"all\"") %>%
  add_input_node(node = "Raw Data", col = "red", ignore = "(download|define|macs2|create|make)") %>%
  change_node_colour("(remove|salmon|counts|analysis)", "red") %>%
  change_node_colour("(download|make|index)", "blue") %>%
  change_node_colour("(find|fastqc|site|template)", "forestgreen") %>%
  str_replace_all("_", "\n") %>%
  str_replace_all("snakemake\ndag", "snakemake_dag") %>%
  str_replace_all("fontsize=[0-9]+", "fontsize=16") %>%
  str_replace_all("(.+graph.+)(\\];)", "\\1, rankdir = LR\\2") %>%
  grViz()
```

## Parameter Description

Parameters set for the complete workflow are:

```{r printYAML, echo=FALSE, results='asis'}
pander(config)
```

## Samples

The samples provided in the config file are:

```{r print-samples, echo=FALSE}
pander(samples)
```


<button type="button" class="btn btn-default btn-sessioninfo" data-toggle="collapse" data-target="#sessioninfo" style="display: block;">
<span class="glyphicon glyphicon-info-sign" aria-hidden="true"></span> Session information
</button>
</p>
<div id="sessioninfo" class="collapse">
```{r session-info, echo=FALSE}
pander::pander(sessionInfo())
```
</div>
